Sound Automated Synthesis of Digital Controllers

### 2016-10-08 paper/paper.tex
expand Figure 2
line breaks

\maketitle
\begin{abstract}
Modern control systems are by and large implemented via digital microcontrollers, 
which are designed to work in an embedded environment, 
within a dynamical plant that can present physical components. 
%
This poses a new dimension to the problem of controllers synthesis,   
which is classically studied in digital control literature. 
%
In this study, we plan to provide a principled approach to extend this literature in a number of directions: 
we are interested in quantitative tasks;  
we focus on automated correct-by-design synthesis algorithms, as opposed to verification objectives; 
and we encompass the complete range of approximations, related to time discretisation, quantisation effects, 
and finite-precision representations (with related rounding errors). To this end, 
we propose a CEGIS-based solution that uses inductive invariant synthesis in conjunction with a digital verification algorithm 
to find a parametric solution for a digital controller given a continuous plant specification. 
\end{abstract}

@@

Modern implementations of embedded control have proliferated with the
availability of low cost devices that can perform simple tasks for
linear control, spanning many areas such as environmental control and
robotic manipulations~\cite{astrom1997computer,Franklin15}. 
The downside to this approach is that in order
to achieve low prices, devices are often limited in their
capabilities, be it a low number of gates in an FPGA or short word
lengths for microcontrollers. In most cases a floating-point
implementation is not feasible and the system can benefit from the
added resolution of a fixed-point representation. The ubiquity of
these devices requires that they have a robust design that ensures the
correct behaviour in a number of different environments.  % For this reason it is
% necessary to provide design tools that formally verify the correctness
% of the digital controller.

@@

%\red{[This section and the next should be merged into a crisp section with the background on DSVerifier, if not included in previous work? 
%Please stick to the introduced notations. Please clarify what the function FWL does, compared to the general approximations introduced above.]}
\red{Alessandro, I've made the changes as requested. Please let me know if this is good enough or if I made it too short}

@@

\begin{mylemma} \cite{astrom1997computer}
%
The pulse transfer function $G(z)$ is a z-domain model that relates the output sampled by a ADC and the digital input provided by ZOH. It is equivalent at the sample times ({\it i.e.}, $kT$ to the continuous transfer function $G(s)$ driven by ZOH).  This pulse transfer
function can be computed as
%
\begin{equation}
\label{eq:pulsetf}
G(z,T) = (1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}

For convenience, we will use only the notation $G(z)$ to represent the pulse transfer function.
\end{mylemma}

Since $G(z)$ is an approximation of $G(s)$ we denote the error in the approximation for time $t \neq kT$ as $\Delta{G}_P$
\begin{align}
\small
\label{parametric_plant_tf}
\hat{G}(z)&=\frac{(b_{0}+\Delta_p b_{0}) +...+(b_{M_G}+\Delta_p b_{M_G})z^{-M_G}}{(a_{0}+\Delta_p a_{0})+...+(a_{N_G}+\Delta_p a_{N_G})z^{-N_G}} \nonumber \\
&=G(z)+\Delta{G}_P(z).
\end{align}
Note that $\Delta{G}_P$ is undefined and upperbound by a user defined percentage such that $\hat{G}(z)$ corresponds to the polynomial set of possible values of $G(z, T): \forall T \in \mathbb{R}$

Let $\tilde C(z)$ and $\tilde G(z)$ be the transfer functions represented using a Fixed Point Finite Word Length with integer size $I$ and fractional size $F$. We shall denote this domain as $\mathbb{R}\langle I,F \rangle$
\begin{align}
\small
\label{digital_controller_plant_tf}
\tilde C(z)&=\frac{\tilde \beta_{0}+\tilde \beta_{1}z^{-1}+...+\tilde \beta_{M_C}z^{-M_C}}{\tilde \alpha_{0}+\tilde \alpha_{1}z^{-1}+...+\tilde \alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
\tilde G(z)&=\frac{\tilde b_{0}+\tilde b_{1}z^{-1}+...+\tilde b_{M_G}z^{-M_G}}{\tilde a_{0}+\tilde a_{1}z^{-1}+...+\tilde a_{N_G}z^{-N_G}}.
\end{align}

Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$ domain, 
 $$\forall i \leq N_C, j \leq M_C\ \  \tilde \beta_{i} \equiv \beta_{i} \wedge \tilde \alpha_{j} \equiv \alpha_{j} \Leftrightarrow \tilde C(z) \equiv C(z)$$
However, given a real plant and real controller $\hat{G}(z), \hat{C}(z)$,
\begin{align}
\small
\label{FWL_tf}
&\exists f (x \in \mathbb{R}) \rightarrow \tilde x \in \mathbb{R}\langle I,F \rangle : \\
\label{digital_controller_tf}
\tilde C(z)&=\frac{(\hat{\beta}_{0}+\Delta_q \hat{\beta}_{0}) +...+(\hat{\beta}_{M_C}+\Delta_q \hat{\beta}_{M_C})z^{-M_C}}{(\hat{\alpha}_{0}+\Delta_q \hat{\alpha}_{0})+...+(\hat{\alpha}_{N_C}+\Delta_q \hat{\alpha}_{N_C})z^{-N_G}} \nonumber \\
&= C(z) =\hat{C}(z) + \Delta{C}(z)\\
\label{digital_plant_tf}
\tilde G(z)&=\frac{(\hat{b}_{0}+\Delta_q \hat{b}_{0}) +...+(\hat{b}_{M_G}+\Delta_q \hat{b}_{M_G})z^{-M_G}}{(\hat{a}_{0}+\Delta_q \hat{a}_{0})+...+(\hat{a}_{N_G}+\Delta_q \hat{a}_{N_G})z^{-N_G}} \nonumber \\
&=G(z)+\Delta{G}_P(z)+\Delta{G}_Q(z).
\end{align}

@@

This intermediate ANSI-C code model $(\tilde C(z),\mathfrak{G}(z),\phi)$
containing a specification $\phi$ for the verification properties ({\it i.e.}, robust stability)
is translated into SAT/SMT formulae by a back-end model-checking tool ({\it e.g.},
CBMC~\cite{ClarkeKL04} or ESBMC~\cite{CordeiroFM12}).
If any property violation is found, then DSVerifier reports a
counterexample, which contains system inputs or parametric deviations
$\Delta{G}$ that lead the system to a failure.  A successful verification
result is reported if the system is safe with respect to $\phi$.  In
particular, the stability verification is complete and sound, since it does
not depend on system inputs and outputs.
In the case of stability, the specification $\phi$ consists in a number of assumptions
on the polynomial coefficients, following Jury's Criteria, as well as the restrictions
set in the representation of these coefficients by the function $f$ (cf. Eq. \eqref{FWL_tf}).

@@

Using an optimisation tool, the work in~\cite{DBLP:conf/hybrid/WangGRJF16} has designed 
a high-performance controller, which is characterised by the following z-domain transfer function:
\begin{equation}
\label{Eq:running-example-controller}
C\left(z\right) := \frac{2.72z^2 - 4.153z + 1.896}{z^2 - 1.844z + 0.8496}.
\end{equation}
The authors claim that this controller $C(z)$ stabilizes the closed-loop system for the discrete plant model $G(z)$.  
However, if implementation aspects are considered during verification ({\it e.g.}, fixed-point arithmetic 
and word-length), then this closed-loop system becomes unstable.
%
Indeed, the actual implementation of $C(z)$ using \param{4}{16} implementation format 
({\it i.e.}, $4$ bits for the integer part and $16$ bits for the fractional part)
can be represented by the following expression: 
\begin{equation}
\label{Eq:running-example-controller-quantized}
\resizebox{0.5\textwidth}{!}{
$C_{q}\left(z\right) {:=} \frac{2.7199859619140625z^2{-}4.1529998779296875z
{+}1.89599609375}{z^2{-}1.843994140625z+0.8495941162109375},$
}
\end{equation} 
which makes the system unstable in closed-loop using fixed-point arithmetic and 
the typical series loop configuration, where $C_{q}\left(z\right)$ and $G\left(z\right)$ 
are in the forward path.

@@

In \cite{widrow1956}, Widrow models the quantization noise as a gaussian model in which
the second moment (variance) of the probability density function (pdf) of the quantization
noise is given by
%
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
where $q$ is the step size of quantizer, thus supporting this assumption.
\end{myassumption}

@@

In order to synthesise closed-loop digital control systems, we use a
program synthesis engine.  Nowadays, such engines are used
increasingly often in program verification
\cite{DBLP:conf/lpar/DavidKL15,DBLP:conf/cav/0001A14}.  Our program
synthesiser makes use of Counter-Example Guided Inductive Synthesis
(CEGIS) \cite{sketch}.  We start by presenting its general
architecture followed by describing the parts specific to closed-loop control
systems.

@@

\paragraph{Synthesis elements specific to control systems}
Next, we describe the specific property that we pass to the 
program synthesiser as the specification $\sigma$.
There are two different algorithms in DSVerifier that can be used for
stability verification~\cite{DBLP:journals/dafes/BessaICF16}, 
one based on Schur's decomposition and another one based on 
Jury's criteria~\cite{astrom1997computer}.
Here, we choose Jury's method~\cite{astrom1997computer}, due to its efficiency.

@@

In recent years the development of program synthesis has found many ways to
obtain correct by design programs from high-level specifications.  One such
approach~\cite{itzhaky2010simple} looks to inductively synthesise invariants
to generate the desired programs.  Since these digital programs are already
constrained by designed to factors such as FWL, these approaches are ideal
for the synthesis of parametric controllers. 
In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS for the
synthesis of switching controllers for stabilizing continuous-time plants
with polynomial dynamics.  The work extends to its application on affine
systems, finding its major challenge in the hardness of solving linear
arithmetic on present SMT solvers.  Since this approach uses switching
states instead of linear dynamics in the digital controller, it entirely
circumvents the FWL problem.  It is also not suitable for the kind of
control we seek to synthesise.  What is needed in this regard is the
combination of a synthesis engine with a control verification tool that
addresses the challenges presented here in the form of FWL effects and
stability measures for LTI SISO controllers.  We take the former
from~\cite{DBLP:conf/lpar/DavidKL15} and the latter from \cite{daes20161} whilst enhancing the
procedure by evaluating the quantaization effects of the Hardware interfaces
(ADC/DAC) to obtain an accurate discrete time FWL representation of the
continuous dynamics.

>>>>

\maketitle

\begin{abstract}
%
Modern control systems are by and large implemented via digital
microcontrollers, which are designed to work in an embedded environment,
within a dynamical plant that can present physical components.
%
This poses a new dimension to the problem of controllers synthesis, which is
classically studied in digital control literature.
%
In this study, we plan to provide a principled approach to extend this
literature in a number of directions: we are interested in quantitative
tasks; we focus on automated correct-by-design synthesis algorithms, as
opposed to verification objectives; and we encompass the complete range of
approximations, related to time discretisation, quantisation effects, and
finite-precision representations (with related rounding errors).  To this
end, we propose a CEGIS-based solution that uses inductive invariant
synthesis in conjunction with a digital verification algorithm to find a
parametric solution for a digital controller given a continuous plant
specification.
%
\end{abstract}

@@

Modern implementations of embedded control have proliferated with the
availability of low cost devices that can perform simple tasks for linear
control, spanning many areas such as environmental control and robotic
manipulations~\cite{astrom1997computer,Franklin15}.  The downside to this
approach is that in order to achieve low prices, devices are often limited
in their capabilities, be it a low number of gates in an FPGA or short word
lengths for microcontrollers.  In most cases a floating-point implementation
is not feasible and the system can benefit from the added resolution of a
fixed-point representation.  The ubiquity of these devices requires that
they have a robust design that ensures the correct behaviour in a number of
different environments.
%
% For this reason it is
% necessary to provide design tools that formally verify the correctness
% of the digital controller.

@@

%\red{[This section and the next should be merged into a crisp section with the background on DSVerifier, if not included in previous work? 
%Please stick to the introduced notations. Please clarify what the function FWL does, compared to the general approximations introduced above.]}

\red{Alessandro, I've made the changes as requested. Please let me know if this is good enough or if I made it too short}

@@

\begin{mylemma}\cite{astrom1997computer}
%
The pulse transfer function $G(z)$ is a z-domain model that relates the
output sampled by a ADC and the digital input provided by ZOH.  It is
equivalent at the sample times ({\it i.e.}, $kT$ to the continuous transfer
function $G(s)$ driven by ZOH).  This pulse transfer function can be
computed as
%
\begin{equation}
\label{eq:pulsetf}
G(z,T) = (1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
%
For convenience, we will use only the notation $G(z)$ to represent the pulse
transfer function.
%
\end{mylemma}

Since $G(z)$ is an approximation of $G(s)$ we denote the error in the
approximation for time $t \neq kT$ as $\Delta{G}_P$
%
\begin{align}
\small
\label{parametric_plant_tf}
\hat{G}(z)&=\frac{(b_{0}+\Delta_p b_{0}) +...+(b_{M_G}+\Delta_p b_{M_G})z^{-M_G}}{(a_{0}+\Delta_p a_{0})+...+(a_{N_G}+\Delta_p a_{N_G})z^{-N_G}} \nonumber \\
&=G(z)+\Delta{G}_P(z).
\end{align}
%
Note that $\Delta{G}_P$ is undefined and upperbound by a user defined
percentage such that $\hat{G}(z)$ corresponds to the polynomial set of
possible values of $G(z, T): \forall T \in \mathbb{R}$

Let $\tilde C(z)$ and $\tilde G(z)$ be the transfer functions represented
using a Fixed Point Finite Word Length with integer size $I$ and fractional
size $F$.  We shall denote this domain as $\mathbb{R}\langle I,F \rangle$
%
\begin{align}
\small
\label{digital_controller_plant_tf}
\tilde C(z)&=\frac{\tilde \beta_{0}+\tilde \beta_{1}z^{-1}+...+\tilde \beta_{M_C}z^{-M_C}}{\tilde \alpha_{0}+\tilde \alpha_{1}z^{-1}+...+\tilde \alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
\tilde G(z)&=\frac{\tilde b_{0}+\tilde b_{1}z^{-1}+...+\tilde b_{M_G}z^{-M_G}}{\tilde a_{0}+\tilde a_{1}z^{-1}+...+\tilde a_{N_G}z^{-N_G}}.
\end{align}

Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$
domain,
%
 $$\forall i \leq N_C, j \leq M_C\ \  \tilde \beta_{i} \equiv \beta_{i} \wedge \tilde \alpha_{j} \equiv \alpha_{j} \Leftrightarrow \tilde C(z) \equiv C(z)$$
%
However, given a real plant and real controller $\hat{G}(z), \hat{C}(z)$,
%
\begin{align}
\small
\label{FWL_tf}
&\exists f (x \in \mathbb{R}) \rightarrow \tilde x \in \mathbb{R}\langle I,F \rangle : \\
\label{digital_controller_tf}
\tilde C(z)&=\frac{(\hat{\beta}_{0}+\Delta_q \hat{\beta}_{0}) +...+(\hat{\beta}_{M_C}+\Delta_q \hat{\beta}_{M_C})z^{-M_C}}{(\hat{\alpha}_{0}+\Delta_q \hat{\alpha}_{0})+...+(\hat{\alpha}_{N_C}+\Delta_q \hat{\alpha}_{N_C})z^{-N_G}} \nonumber \\
&= C(z) =\hat{C}(z) + \Delta{C}(z)\\
\label{digital_plant_tf}
\tilde G(z)&=\frac{(\hat{b}_{0}+\Delta_q \hat{b}_{0}) +...+(\hat{b}_{M_G}+\Delta_q \hat{b}_{M_G})z^{-M_G}}{(\hat{a}_{0}+\Delta_q \hat{a}_{0})+...+(\hat{a}_{N_G}+\Delta_q \hat{a}_{N_G})z^{-N_G}} \nonumber \\
&=G(z)+\Delta{G}_P(z)+\Delta{G}_Q(z).
\end{align}

@@

This intermediate ANSI-C code model $(\tilde C(z),\mathfrak{G}(z),\phi)$
containing a specification $\phi$ for the verification properties ({\it
i.e.}, robust stability) is translated into SAT/SMT formulae by a back-end
model-checking tool ({\it e.g.}, CBMC~\cite{ClarkeKL04} or
ESBMC~\cite{CordeiroFM12}).  If any property violation is found, then
DSVerifier reports a counterexample, which contains system inputs or
parametric deviations $\Delta{G}$ that lead the system to a failure.  A
successful verification result is reported if the system is safe with
respect to $\phi$.  In particular, the stability verification is complete
and sound, since it does not depend on system inputs and outputs.  In the
case of stability, the specification $\phi$ consists in a number of
assumptions on the polynomial coefficients, following Jury's Criteria, as
well as the restrictions set in the representation of these coefficients by
the function $f$ (cf.~Eq.~\eqref{FWL_tf}).

@@

Using an optimisation tool, the work in~\cite{DBLP:conf/hybrid/WangGRJF16}
has designed a high-performance controller, which is characterised by the
following z-domain transfer function:
%
\begin{equation}
\label{Eq:running-example-controller}
C\left(z\right) := \frac{2.72z^2 - 4.153z + 1.896}{z^2 - 1.844z + 0.8496}.
\end{equation}
%
The authors claim that this controller $C(z)$ stabilizes the closed-loop
system for the discrete plant model $G(z)$.  However, if implementation
aspects are considered during verification ({\it e.g.}, fixed-point
arithmetic and word-length), then this closed-loop system becomes unstable.
%
Indeed, the actual implementation of $C(z)$ using \param{4}{16}
implementation format ({\it i.e.}, $4$ bits for the integer part and $16$
bits for the fractional part) can be represented by the following
expression:
%
\begin{equation}
\label{Eq:running-example-controller-quantized}
\resizebox{0.5\textwidth}{!}{
$C_{q}\left(z\right) {:=} \frac{2.7199859619140625z^2{-}4.1529998779296875z
{+}1.89599609375}{z^2{-}1.843994140625z+0.8495941162109375},$
}
\end{equation} 
%
which makes the system unstable in closed-loop using fixed-point arithmetic
and the typical series loop configuration, where $C_{q}\left(z\right)$ and
$G\left(z\right)$ are in the forward path.

@@

In~\cite{widrow1956}, Widrow models the quantization noise as a gaussian
model in which the second moment (variance) of the probability density
function (pdf) of the quantization noise is given by
%
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
where $q$ is the step size of quantizer, thus supporting this assumption.
\end{myassumption}

@@

In order to synthesise closed-loop digital control systems, we use a program
synthesis engine.  Nowadays, such engines are used increasingly often in
program verification~\cite{DBLP:conf/lpar/DavidKL15,DBLP:conf/cav/0001A14}. 
Our program synthesiser makes use of Counter-Example Guided Inductive
Synthesis (CEGIS) \cite{sketch}.  We start by presenting its general
architecture followed by describing the parts specific to closed-loop
control systems.

@@

\paragraph{Synthesis elements specific to control systems}

Next, we describe the specific property that we pass to the program
synthesiser as the specification $\sigma$.  There are two different
algorithms in DSVerifier that can be used for stability
verification~\cite{DBLP:journals/dafes/BessaICF16}, one based on Schur's
decomposition and another one based on Jury's
criteria~\cite{astrom1997computer}.  Here, we choose Jury's
method~\cite{astrom1997computer}, due to its efficiency.

@@

In recent years the development of program synthesis has found many ways to
obtain correct by design programs from high-level specifications.  One such
approach~\cite{itzhaky2010simple} looks to inductively synthesise invariants
to generate the desired programs.  Since these digital programs are already
constrained by designed to factors such as FWL, these approaches are ideal
for the synthesis of parametric controllers. 
In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS for the
synthesis of switching controllers for stabilizing continuous-time plants
with polynomial dynamics.  The work extends to its application on affine
systems, finding its major challenge in the hardness of solving linear
arithmetic on present SMT solvers.  Since this approach uses switching
states instead of linear dynamics in the digital controller, it entirely
circumvents the FWL problem.  It is also not suitable for the kind of
control we seek to synthesise.  What is needed in this regard is the
combination of a synthesis engine with a control verification tool that
addresses the challenges presented here in the form of FWL effects and
stability measures for LTI SISO controllers.  We take the former
from~\cite{DBLP:conf/lpar/DavidKL15} and the latter from~\cite{daes20161}
whilst enhancing the procedure by evaluating the quantaization effects of
the Hardware interfaces (ADC/DAC) to obtain an accurate discrete time FWL
representation of the continuous dynamics.
